---
author: "Raghu Ramnath"
date: "3/11/2017"
title: "Raghu_606_Project_Final"
output: 
  html_document:
    code_folding: show
    css: ~/Lab7/lab.css
    fig_height: 7
    fig_width: 10
    highlight: zenburn
    theme: cerulean
    toc: yes
    toc_float: yes
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```     


## Part I - Introduction

  In this Project, I will analyze the shooting incidents in all the 50 States in US based on the data from year 2013 to 2015. The objective is to find the States having highest shooting rate and people killed and wounded and see if there is any trend by comparison of the count of incidents. Has the incidents raised or lowered in the three years. Is there any particular month having high incident rate in any state. This would help to understand if any measure has been taken already or need to take to lower the incidents. This would also warn in terms of understanding high crime prone State.
    
### Environment Prep

```{r, warning=FALSE, message=FALSE}

if (!require('dplyr')) install.packages('dplyr')
if (!require('stringr')) install.packages('stringr')
if (!require('ggplot2')) install.packages('ggplot2')
if (!require('lubridate')) install.packages('lubridate')
if (!require('tidyr')) install.packages('tidyr')
if (!require('doBy')) install.packages('doBy')
if (!require('knitr')) install.packages('knitr')
if (!require('reshape')) install.packages('reshape')
if (!require('lattice')) install.packages('lattice')

```     


## Part 2 - Data

```{r D1, include=TRUE , warning=FALSE, message=FALSE}

yr_2013 <- read.csv(file = ("https://raw.githubusercontent.com/raghu74us/606/master/2013_shooter.csv"),sep =",")

yr_2014 <- read.csv(file = ("https://raw.githubusercontent.com/BuzzFeedNews/2015-12-mass-shooting-intervals/master/data/2014MASTER.csv"),sep =",")

yr_2015 <- read.csv(file = ("https://raw.githubusercontent.com/BuzzFeedNews/2015-12-mass-shooting-intervals/master/data/2015CURRENT.csv"),sep =",")

```     


### Cases

  There are only 3 years of data available which totals to 1050 observations and 9 variables. This data set has the year, month, City, state and incident date and total killed and wounded.

### Describe the method of data collection.

  Data is available in Buzfeed News. There are data files for each year from 2013 to 2015.


### Type of study

  This study is observational.

### Data Source

  Data is available under the below link.
https://github.com/BuzzFeedNews/2015-12-mass-shooting-intervals/tree/master/data

### Response

  Response variable is the number of  incidents in a month in a city, State. It is numerical. 


### Explanatory

  Number of incidents each month. Has the incidents happened each month what is the count of it.
  This need to be done for atleast 3 years to decide which city has highest incident and the reoccurrence  chances each month. 


## Part 3 - Exploratory data analysis

```{r s1, include=TRUE}

#not including the columns that are not needed for analysis.
yr_2013 <- yr_2013[ , c('date', 'killed', 'wounded', 'location')]
colnames(yr_2013) <- c('Date', 'Dead', 'Injured', 'Location')
yr_2013$Year <- "2013"
yr_2014 <- yr_2014[ , c('Date', 'Dead', 'Injured', 'Location')]
yr_2014$Year <- "2014"
yr_2015 <- yr_2015[ , c('Date', 'Dead', 'Injured', 'Location')]
yr_2015$Year <- "2015"

sh <-rbind(yr_2013, yr_2014, yr_2015)


     #split state and city based on location
     str(sh)
     head(sh,5)   
   
     cit_st<-str_split_fixed(sh$Location, ",", 2)
     sh   <- cbind(sh, cit_st)
     sh$Year<- as.factor(sh$Year)
     colnames(sh) <- c('date', 'killed', 'wounded', 'location', "Year","City", "State")
     sh$State<- as.factor(sh$State)
     #Add month and year based on date.
     sh$date <- mdy(sh$date)
     
     str(sh) 
     
     #y1<-year(sh$date)
     m1<-as.character(sh$date, format="%b")
     d1<-as.character(sh$date, format="%d")
     
     df1  <- cbind(d1,m1)
     colnames(df1)  <- c('day','month')
     
     sh <- cbind(sh , df1)
     
     sh$State <- trimws(toupper(sh$State))
     sh$State <- replace(sh$State, sh$State=="ILLINOIS", "IL")
     sh$State <- replace(sh$State, sh$State=="TENNESSEE", "TN")
     sh$State <- replace(sh$State, sh$State=="KANSAS", "KS")
     sh$State <- replace(sh$State, sh$State=="LOUISIANA", "MO")
     sh$State <- replace(sh$State, sh$State=="OHIO", "OH")
     sh$State <- replace(sh$State, sh$State=="D.C.", "DC")
     sh$State <- replace(sh$State, sh$State=="PUERTO RICO", "NA")
     sh$State <- replace(sh$State, sh$State=="", "NA")
          
     
     sh_us_sum <- sh %>%  
         select(Year,State,killed,wounded) %>%
          group_by(Year,State )  %>%
            summarise( total_incidents= n() ) %>%
               top_n(5)
     
     
     ggplot(sh_us_sum, aes(State,total_incidents)) + 
    geom_line(aes(color=Year, group=Year)) +
     ggtitle("Shooting in Top 5 States from 2013 to 2015")    
       
     sh1 <- sh[ !((trimws(sh$State)) != "CA"), ]
    
    #summarize by total killed and wounded in California  
     sh_kw_ca <- sh1  %>%
        select(Year,State,month,killed,wounded) %>%
          group_by(Year,State,month )  %>%
           summarise_each(funs(sum) ) %>%
             arrange(Year,State,month)
     
     sh_kw_ca$Tot_kw<- sh_kw_ca$killed + sh_kw_ca$wounded
     
    #summarize by total incidents in California        
      sh_inc_ca <- sh1  %>%
        select(Year,State,month,killed+wounded) %>%
          group_by(Year,State,month )  %>%
           summarise( total_incidents= n() ) 
      
      
knitr::opts_chunk$set(echo = TRUE)
```

### Plots

```{r, P1, include = TRUE}
#Incidents from 2013 to 2015 in California Monthwise.      
ggplot( sh_inc_ca , aes(sh_inc_ca$month, 
                sh_inc_ca$total_incidents, group=Year, colour=Year)) +
  geom_line() +
  geom_point() +
  labs(x="Month", colour="Year") +
  theme_classic() +
     ggtitle("Incidents from 2013 to 2015 in California Monthwise")

#Total Killed and Wounded from 2013 to 2015 in California Monthwise.      
ggplot( sh_kw_ca , aes( sh_kw_ca$month,sh_kw_ca$Tot_kw, group=Year, colour=Year))  +
  geom_line() +
  geom_point() +
  labs(x="Month", colour="Year") +
  theme_classic() +
    ggtitle("Total killed and Wounded from 2013 to 2015 in California Monthwise")


knitr::opts_chunk$set(echo = TRUE)
```     

## Part 4 - Inference

### Hypothesis

$H_0: \mu_{TJan} = \mu_{Feb} =... \mu_{Dec} $.  The avg monthly shooting does not vary across all year.

$H_A:$ The avg monthly shooting across some or all months does vary.

### Check Conditions

```{r s2, include=TRUE}

#box plot from Jan to Dec in CA from 2013-15

bwplot(sh_inc_ca$total_incidents ~ sh_inc_ca$month | sh_inc_ca$Year, sh_inc_ca,main="Incidents by Month from 2013-15",xlab="month",ylab="Total Incidents")

#wide format
      sh_inc_w <- sh_inc_ca  %>%
                   spread(month, total_incidents, fill=NA, convert = FALSE, drop=TRUE, sep=NULL)

      summary(sh_inc_w)
      boxplot(sh_inc_w[3:14], main="Incidents by Month from 2013-15",xlab="month",ylab="Total Incidents")

# Average mean and SD of Jan to Dec.
      
sh_inc_ca_msc <- data.frame(State=sh_inc_ca$State,Month= sh_inc_ca$month,Incidents= sh_inc_ca$total_incidents)

sh_inc_ca_TAS <- aggregate(Incidents ~ Month, data = sh_inc_ca_msc, function(x) c(Total = sum(x), Average = mean(x), SD = sd(x) )) 

sh_inc_ca_TAS<- cbind(sh_inc_ca_TAS[-ncol(sh_inc_ca_TAS)], sh_inc_ca_TAS[[ncol(sh_inc_ca_TAS)]])

sh_inc_ca_TAS

p<- ggplot(sh_inc_ca_TAS, aes(x=sh_inc_ca_TAS$Total, y=sh_inc_ca_TAS$Average, group=sh_inc_ca_TAS$Month, color=sh_inc_ca_TAS$Month)) + 
  geom_line() +
  geom_point()+
  geom_errorbar(aes(ymin=sh_inc_ca_TAS$Total-sh_inc_ca_TAS$SD, ymax=sh_inc_ca_TAS$Total+sh_inc_ca_TAS$SD), width=.2,
                 position=position_dodge(0.05))
print(p)
# Finished line plot
p + labs(title="Shooting Incidents by Month in California", x="Incidents Total", y = "Average by Month")+
   theme_classic() 


knitr::opts_chunk$set(echo = TRUE)

```

We believe that the shooting data from each month is independent and varies across all months.

## Part 5 - Conclusion

The study concludes that the shooting incidents has been decreasing from 2013 to 2015. The number of incidents
occurring each month is not related or dependent on previous months or years.